---
title: "Mini Project 1"
author: "Michaela Gustafson"
date: "10/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spatial Data in R: Mini Project 1

The goals for this phase of your final project are:

Articulate an interesting research question based on a dataset you’d like to learn more about.

Develop a spatial database that contains potentially relevant explanatory variables that you’d like to explore in the context of that research question.

Demonstrate an understanding of the various workflow elements involved in designing and constructing a spatial database for subsequent visualization and analysis.

```{r library}
library(here)
library(tidyr)
library(dplyr)
library(terra)
library(sf)
```

# Upload zombie data

```{r zombieload}
here()
zom <- read.csv(here("zombies.csv"))

head(zom); dim(zom)

zombie.csv <- read.csv(here("New_Zealand_Zombie_Attack_Data.csv"))

zombie.shp <- read_sf(here("New_Zealand_Zombie_Attack_Data.shp"))
plot(zombie.shp)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
